---
title: "MOFA"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(tidyr)
library(MOFA)
library(MOFAdata)
library(reticulate)
use_python("/Users/senosam/opt/anaconda3/bin/python", required=TRUE) 
```


# Load MOFA models
```{r, echo=FALSE}
load('/Users/senosam/Documents/Massion_lab/data_integration/MOFA_model.RData')
```


## MOFA model + xCell + VIPER 25
```{r}
plotDataOverview(MOFA_m_x25)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_x25)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_x25)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_x25, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_prcnt", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_x25, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_x25, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_x25, 
  view = "RNA_xcell", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_xcell'
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_xcell", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_xcell", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_xcell", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_xcell", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_x25, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="RNA_viper25", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x25, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_x25,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r}
plotFactorScatter(
  MOFA_m_x25,
  factors = 1:2,
  color_by = "ImmuneScore"      # color by the IGHV values that are part of the training data
  #shape_by = "trisomy12"  # shape by the trisomy12 values that are part of the training data
)
```



```{r}
plotFactorScatters(
  MOFA_m_x25,
  factors = c(1:5),
  color_by = "ImmuneScore"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x25,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x25,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```


```{r}
plotFactorScatters(
  MOFA_m_x25,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x25,
  factors = c(1:5),
  color_by = "EGFR"
)
```


## MOFA model + xCell + VIPER 4
```{r}
plotDataOverview(MOFA_m_x4)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_x4)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_x4)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_x4, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_prcnt", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_x4, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_x4, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_x4, 
  view = "RNA_xcell", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_xcell'
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_xcell", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_xcell", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_xcell", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="RNA_xcell", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_x4, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_x4, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_x4,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_x4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r}
plotFactorScatter(
  MOFA_m_x4,
  factors = 1:2,
  color_by = "ImmuneScore"      # color by the IGHV values that are part of the training data
  #shape_by = "trisomy12"  # shape by the trisomy12 values that are part of the training data
)
```



```{r}
plotFactorScatters(
  MOFA_m_x4,
  factors = c(1:5),
  color_by = "ImmuneScore"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x4,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x4,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x4,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_x4,
  factors = c(1:5),
  color_by = "EGFR"
)
```


## MOFA model + EpiDISH + VIPER 25
```{r}
plotDataOverview(MOFA_m_e25)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_e25)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_e25)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_e25, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_prcnt", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_e25, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_e25, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_e25, 
  view = "RNA_epidish", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_epidish'
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_epidish", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_epidish", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_epidish", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="RNA_epidish", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_e25, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e25, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_e25,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e25,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```




```{r}
plotFactorScatters(
  MOFA_m_e25,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e25,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e25,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e25,
  factors = c(1:5),
  color_by = "EGFR"
)
```

## MOFA model + EpiDISH + VIPER 4
```{r}
plotDataOverview(MOFA_m_e4)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_e4)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_e4)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_epidish", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_epidish'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_e4,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "EGFR"
)
```


## MOFA model SIMPLE ()
```{r}
plotDataOverview(MOFA_m_e4)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_e4)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_e4)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_epidish", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_epidish'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_e4,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "EGFR"
)
```


## MOFA model SIMPLE + RADIOMICS
```{r}
plotDataOverview(MOFA_m_e4)
```

### Analyze trained MOFA model
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFA_m_e4)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFA_m_e4)
```

### Characterization of individual factors

#### CyTOF percentages
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_prcnt", 
  factor=2
)
```


#### CyTOF expression
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=3
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=5
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="CyTOF_exp", 
  factor=2
)
```

#### RNA Seq Top genes
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_topgenes", 
  factor=1
)
```

#### RNA Seq Deconvolution
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "RNA_epidish", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_epidish'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=2
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=4
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="RNA_epidish", 
  factor=5
)
```

#### WES
```{r}
plotWeightsHeatmap(
  MOFA_m_e4, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

```{r}
plotTopWeights(
  MOFA_m_e4, 
  view="mutation", 
  factor=1
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFA_m_e4,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 1:2

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 3:4

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r fig.width = 9, fig.height = 5}
interestingFactors <- 5:6

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFA_m_e4,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Th_cells"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "Tc_cells"
)
```


```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "KRAS"
)
```

```{r}
plotFactorScatters(
  MOFA_m_e4,
  factors = c(1:5),
  color_by = "EGFR"
)
```


