---
title: "Summary of RF"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyverse)
library(MOFA)
library(MOFAdata)
library(reticulate)
library(reticulate)
use_python("/Users/senosam/opt/anaconda3/bin/python", required=TRUE) 
```

# Example data set
```{r}
data <- makeExampleData()
class(data)
class(data$view_1)
data$view_1[1:5, 1:5]
data$view_2[1:5, 1:5]
data$view_3[1:5, 1:5]
summary(data$view_1[,1:10])
```

```{r}
set.seed(1234)
MOFAobject <- createMOFAobject(data)
```

```{r}
plotDataOverview(MOFAobject)
```

## Prepare MOFA: Set the training and model options
```{r}
TrainOptions <- getDefaultTrainOptions()
ModelOptions <- getDefaultModelOptions(MOFAobject)
DataOptions <- getDefaultDataOptions()

TrainOptions$DropFactorThreshold <- 0.01
```

## Run MOFA
```{r}
n_inits <- 3
MOFAlist <- lapply(seq_len(n_inits), function(it) {
  
  TrainOptions$seed <- 2018 + it
  
  MOFAobject <- prepareMOFA(MOFAobject,
                            DataOptions = DataOptions,
                            ModelOptions = ModelOptions,
                            TrainOptions = TrainOptions
)
  
  runMOFA(MOFAobject)
})
```

## Compare different random inits and select the best model
```{r}
compareModels(MOFAlist)
```
With compareFactors we can get an overview of how robust the factors are between different model instances.
```{r}
compareFactors(MOFAlist)
```

For down-stream analyses we recommned to choose the model with the best ELBO value as is done by selectModel.
```{r}
MOFAobject <- selectModel(MOFAlist, plotit = FALSE)
MOFAobject
```


## Downstream analyses
On the trained MOFAobject we can now start looking into the inferred factors, its weights etc. Here the data was generated using five factors, whose activity patterns we can recover using plotVarianceExplained.
```{r}
plotVarianceExplained(MOFAobject)
```




# Mafe's data (CyTOF, RNA Seq, WES)
```{r, echo=FALSE}
load('/Users/senosam/Documents/Massion_lab/data_integration/MO_data.Rdata')
e <- c(15,22,25,33,36,41,52,56,61,64,68,80,83,94,103,106,111,124,127,131)
CyTOF_exp <- CyTOF_exp[-e,]
rownames(RNA_topclust_E) <- sapply(strsplit(rownames(RNA_topclust_E), "\\."), "[[", 1)
rownames(RNA_top12K_E) <- sapply(strsplit(rownames(RNA_top12K_E), "\\."), "[[", 1)
```


## Preprocessing data
```{r}
mofa_data <- list('CyTOF_exp'=CyTOF_exp, 
                  'CyTOF_prcnt'=CyTOF_prcnt, 
                  'RNA_topgenes'=RNA_top12K_E, 
                  'RNA_xcell'=RNA_xcell,
                  'mutation'=mut_dt)
set.seed(1234)
MOFAobject <- createMOFAobject(mofa_data)
```

```{r}
plotDataOverview(MOFAobject)
```

## Prepare MOFA: Set the training and model options
The most important options the user needs to define are:

scaleViews: logical indicating whether to scale views to have unit variance. As long as the scale of the different data sets is not too high, this is not required. Default is FALSE.

removeIncompleteSamples: logical indicating whether to remove samples that are not profiled in all omics. The model can cope with missing assays, so this option is not required. Default is FALSE.

### Data options
```{r}
DataOptions <- getDefaultDataOptions()
DataOptions 
```

### Model options
Next, we define model options. The most important are:

numFactors: number of factors (default is 0.5 times the number of samples). By default, the model will only remove a factor if it explains exactly zero variance in the data. You can increase this threshold on minimum variance explained by setting TrainOptions$dropFactorThreshold to a value higher than zero.

likelihoods: likelihood for each view. Usually we recommend gaussian for continuous data, bernoulli for binary data and poisson for count data. By default, the model tries to guess it from the data.

sparsity: do you want to use sparsity? This makes the interpretation easier so it is recommended (Default is TRUE).
```{r}
ModelOptions <- getDefaultModelOptions(MOFAobject)
ModelOptions
```

### Training options
Next, we define training options. The most important are:

maxiter: maximum number of iterations. Ideally set it large enough and use the convergence criterion TrainOptions$tolerance.

tolerance: convergence threshold based on change in the evidence lower bound. For an exploratory run you can use a value between 1.0 and 0.1, but for a “final” model we recommend a value of 0.01.

DropFactorThreshold: hyperparameter to automatically learn the number of factors based on a minimum variance explained criteria. Factors explaining less than DropFactorThreshold fraction of variation in all views will be removed. For example, a value of 0.01 means that factors that explain less than 1% of variance in all views will be discarded. By default this it zero, meaning that all factors are kept unless they explain no variance at all.
```{r}
TrainOptions <- getDefaultTrainOptions()

# Automatically drop factors that explain less than 2% of variance in all omics
TrainOptions$DropFactorThreshold <- 0.02

TrainOptions$seed <- 2017

TrainOptions
```

## Prepare MOFA
prepareMOFA internally performs a set of sanity checks and fills the DataOptions, TrainOptions and ModelOptions slots of the MOFAobject

```{r}
MOFAobject <- prepareMOFA(
  MOFAobject, 
  DataOptions = DataOptions,
  ModelOptions = ModelOptions,
  TrainOptions = TrainOptions
)
```

## Run MOFA
```{r}
MOFAobject <- runMOFA(MOFAobject)
```

## Analyse trained MOFA model
Disentangling the heterogeneity: calculation of variance explained by each factor in each view
```{r}
# Calculate the variance explained (R2) per factor in each view 
r2 <- calculateVarianceExplained(MOFAobject)
r2$R2Total
# Variance explained by each factor in each view
head(r2$R2PerFactor)
# Plot it
plotVarianceExplained(MOFAobject)
```

### Characterization of individual factors

To get an overview of the weights across all factors in a given view you can use the plotWeightsHeatmap function.


```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "CyTOF_prcnt", 
  factors = 1:5,
  show_colnames = T, main = 'CyTOF_prcnt'
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="CyTOF_prcnt", 
  factor=4
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "CyTOF_exp", 
  factors = 1:5,
  show_colnames = FALSE, main = 'CyTOF_exp'
)
```


```{r}
plotTopWeights(
  MOFAobject, 
  view="CyTOF_exp", 
  factor=1
)
```


```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "RNA_topgenes", 
  factors = 1:5,
  show_colnames = FALSE, main = 'RNA_topgenes'
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="RNA_topgenes", 
  factor=3
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "RNA_xcell", 
  factors = 1:5,
  show_colnames = T, main = 'RNA_xcell'
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="RNA_xcell", 
  factor=2
)
```

```{r}
plotWeightsHeatmap(
  MOFAobject, 
  view = "mutation", 
  factors = 1:5,
  show_colnames = F, main = 'Mutation Data'
)
```

```{r}
plotTopWeights(
  MOFAobject, 
  view="mutation", 
  factor=3
)
```

### Feature set enrichment analysis in the active views
```{r}
# Load reactome annotations
data("reactomeGS") # binary matrix with feature sets in rows and features in columns

# perform enrichment analysis
gsea <- runEnrichmentAnalysis(
  MOFAobject,
  view = "RNA_topgenes",
  feature.sets = reactomeGS,
  alpha = 0.01
)
```

```{r}
plotEnrichmentBars(gsea, alpha=0.01)
```

```{r}
interestingFactors <- 1:5

fseaplots <- lapply(interestingFactors, function(factor) {
  plotEnrichment(
    MOFAobject,
    gsea,
    factor = factor,
    alpha = 0.01,
    max.pathways = 10 # The top number of pathways to display
  )
})

cowplot::plot_grid(fseaplots[[1]], fseaplots[[2]],
                   ncol = 1, labels = paste("Factor", interestingFactors))
```

```{r}
plotFactorScatter(
  MOFAobject,
  factors = 1:2,
  color_by = "ImmuneScore"      # color by the IGHV values that are part of the training data
  #shape_by = "trisomy12"  # shape by the trisomy12 values that are part of the training data
)
```

```{r}
plotFactorScatters(
  MOFAobject,
  factors = 1:3,
  color_by = "ImmuneScore"
)
```

```{r}
plotFactorScatters(
  MOFAobject,
  factors = 3:5,
  color_by = "ImmuneScore"
)
```

```{r}
plotFactorScatters(
  MOFAobject,
  factors = c(1, 3:5),
  color_by = "ImmuneScore"
)
```

```{r}
```

```{r}
```