---
title: "Summary of TMA36 cohort: clinical characteristics"
author: "Mafe Senosain"
date: "`r Sys.Date()`"
output:
    html_document:
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(reshape2)
library(jcolors)
library(ComplexHeatmap)
library(dplyr)
library(tidyr)
library(ggrepel)
library(forcats)
library(ggsignif)
library(ggpubr)
knitr::opts_chunk$set(echo = FALSE)
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/20_ClustAnnot_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/30_DA_functions.R")
source("/Users/senosam/Documents/Repositories/Research/data_analysis_cytof/R/40_DE_functions.R")
```

```{r, echo=FALSE}
# read edited CSV file as CDE_TMA36
CDE <- read.csv(file = '/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2021JAN13_SA.csv')
```

```{r, echo=FALSE}
pie_chart <- function(dt, col_name, plot_title){
    x <- data.frame(table(dt[col_name]))
    colnames(x) <- c('Group', 'value')
    x <- x %>% 
        arrange(desc(value)) #%>%
        #mutate(prop = label_percent(value / sum(value)))

    print(ggplot(x, aes(x = "", y = value, fill = fct_inorder(Group))) +
        geom_bar(width = 1, stat = "identity") +
        coord_polar("y", start = 0) +
        geom_label_repel(aes(label = value), size=5, show.legend = F, nudge_x = 1) +
        #geom_label_repel(aes(label = prop), size=5, show.legend = F, nudge_x = 1) + #%
        guides(fill = guide_legend(title = "Group")) +
        labs(title=plot_title) +
        theme(plot.title = element_text(hjust = 0.5, size=22)))
}

```

# Clinical characteristics distributions on TMA36 cohort
```{r, echo=FALSE}
hist(CDE$SILA, xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE, 'CANARY', 'CANARY')
pie_chart(CDE, 'Stages_simplified', 'Stage')
pie_chart(CDE, 'Hist_predominant', 'Histology')
pie_chart(CDE, 'Death_st', 'Death Status')
pie_chart(CDE, 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE, 'Progression_st', 'Progression Status')
pie_chart(CDE, 'Smoking_Status', 'Smoking Status')
```

# Clinical characteristics distributions by CANARY

## Good prognosis (Indolent)
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'G')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')

```

## Intermediate prognosis
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'I')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')
```

## Poor prognosis (Aggressive)
```{r, echo=FALSE}
k <- which(CDE$CANARY == 'P')
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
pie_chart(CDE[k,], 'Smoking_Status', 'Smoking Status')
```

# Clinical characteristics distributions by SILA Score

```{r, echo=FALSE}
#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/
bp_f <- function(dt, chr_colnm, sila_colnm='SILA', main_title, x_title) {
  dt <- dt[,c(sila_colnm,chr_colnm)]
  dt <- melt(dt)
  colnames(dt)[1]<- 'condition'
  
  x <- compare_means(value ~ condition, dt, method = "wilcox.test", paired = FALSE,
                     group.by = NULL, ref.group = NULL)
  idx <- which(x$p.signif != 'ns')
  ls <- list()
  n=1
  for (i in idx) {
    ls[[n]] <- c(x$group1[idx[n]], x$group2[idx[n]])
    n=n+1
  }
  #print(data.frame(x))
  
  p <- ggplot(dt, aes(x=condition, y=value, fill=condition)) +
    geom_boxplot() +
    labs(title=main_title,x=x_title, y = "SILA score") +
    theme(plot.title = element_text(hjust = 0.5, size=20)) +
    ggsignif::geom_signif(comparisons = ls, 
       map_signif_level=TRUE)
    geom_signif(comparisons = ls, 
              map_signif_level=TRUE)
  return(p)
}

bp_f(CDE, chr_colnm='CANARY', sila_colnm='SILA', main_title='SILA score by CANARY', x_title='CANARY')
bp_f(CDE, chr_colnm='Stages_simplified', sila_colnm='SILA', main_title='SILA score by Stage', x_title='Tumor Stage')
bp_f(CDE, chr_colnm='Hist_predominant', sila_colnm='SILA', main_title='SILA score by Histology', x_title='Predominant Histology')
bp_f(CDE, chr_colnm='Death_st', sila_colnm='SILA', main_title='SILA score by Death Status', x_title='Death Status')
bp_f(CDE, chr_colnm='Recurrence_st', sila_colnm='SILA', main_title='SILA score by Recurrence Status', x_title='Recurrence Status')
bp_f(CDE, chr_colnm='Progression_st', sila_colnm='SILA', main_title='SILA score by Progression Status', x_title='Progression Status')
bp_f(CDE, chr_colnm='Smoking_Status', sila_colnm='SILA', main_title='SILA score by Smoking Status', x_title='Smoking Status')


```

```{r}
# ggpubr::ggscatter(CDE, x = 'SILA', y = c("Age_at_collection"),
#           add = "reg.line", conf.int = TRUE, combine = TRUE,
#           cor.coef = TRUE, cor.method = "spearman", add.params = list(color = 'grey'),
#           xlab = "SILA score", ylab = 'condition')
```


# New classification 

## Identifying outliers
```{r, echo=FALSE}
flag = c()
for (i in 1:nrow(CDE)){
    # if SILA >=0.5 stage shouldn't be 0, histology shouldn't be lepidic
    if (CDE$SILA[i]>=0.5) {
        if( CDE$Stages_simplified[i] %in% c('Stage 0') || 
            CDE$Hist_predominant[i] %in% c('lepidic')) {
            #cat('flag!: ', CDE$pt_ID[i], '\n')
            flag = c(flag, i)
        }
    }
    # if SILA <0.5 stage shouldn't be 3 or 4, histology shouldn't be solid or micropapillary, 
    # patient shouldn't be dead, have a recurrence or a progression, CANARY shouldn't be Poor
    if (CDE$SILA[i]<0.5) {
        if ( CDE$Stages_simplified[i] %in% c('Stage III', 'Stage IV') || 
            CDE$Hist_predominant[i] %in% c('solid', 'micropapillary') || 
            CDE$Death_st[i] == 'Yes' || 
            CDE$Recurrence_st[i] == 'Yes' || 
            CDE$Progression_st[i] == 'Yes' ||
            CDE$CANARY[i] == 'P' ){
            #cat('flag!: ', CDE$pt_ID[i], '\n')
            flag = c(flag, i)
        }

    }
}



fl_df <- CDE[flag, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]

fl_g <- fl_df[which(fl_df$CANARY == 'G'),]
fl_i <- fl_df[which(fl_df$CANARY == 'I'),]
fl_p <- fl_df[which(fl_df$CANARY == 'P'),]
fl_df <- rbind(fl_g, fl_i, fl_p)
knitr::kable(fl_df)
```

## Option 1: Indolent < 0.5 | Aggressive  >= 0.5

```{r, echo=FALSE}
ind <- CDE[which(CDE$SILA <= 0.5),]
agg <- CDE[which(CDE$SILA > 0.5),]
n_op1 <- c(rep(0, nrow(CDE)))
```

### Indolent as SILA < 0.5
```{r, echo=FALSE}
k <- which(CDE$SILA < 0.5)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')

i_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(i_df)
n_op1[k] <- 'ind'
```

### AGGRESSIVE as SILA >= 0.5
```{r, echo=FALSE}
k <- which(CDE$SILA >= 0.5)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')

a_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(a_df)
n_op1[k] <- 'agg'
```

## Option 2: Indolent < 0.4 | intermediate > 0.4 <= 0.6 | Aggressive  > 0.6

```{r, echo=FALSE}
ind <- CDE[which(CDE$SILA <= 0.4),]
int <- CDE[which(CDE$SILA > 0.4 & CDE$SILA <= 0.6),]
agg <- CDE[which(CDE$SILA < 0.6),]
n_op2 <- c(rep(0, nrow(CDE)))
```

### Indolent as SILA <= 0.4
```{r, echo=FALSE}
k <- which(CDE$SILA <= 0.4)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
i_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(i_df)
n_op2[k] <- 'ind'
```

### Intermediate as SILA > 0.4 <= 0.6
```{r, echo=FALSE}
k <- which(CDE$SILA > 0.4 & CDE$SILA <= 0.6)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
int_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(int_df)
n_op2[k] <- 'int'
```

### AGGRESSIVE as SILA > 0.6
```{r, echo=FALSE}
k <- which(CDE$SILA > 0.6)
hist(CDE$SILA[k], xlab="SILA score", main = 'SILA Score', xlim=c(0,1), col="chocolate", border="brown", las=1, xaxt = "n")
axis(1, at = seq(0, 1, .1))
pie_chart(CDE[k,], 'Stages_simplified', 'Stage')
pie_chart(CDE[k,], 'Hist_predominant', 'Histology')
pie_chart(CDE[k,], 'Death_st', 'Death Status')
pie_chart(CDE[k,], 'Recurrence_st', 'Recurrence Status')
pie_chart(CDE[k,], 'Progression_st', 'Progression Status')
a_df <- CDE[k, c('pt_ID','SILA', 'CANARY', 'Stages_simplified', 'X8th_ed_path_stage', 
    'Hist_predominant', 'Hist_other_patterns', 'Death_st', 'Recurrence_st', 
    'Progression_st')]
knitr::kable(a_df)
n_op2[k] <- 'agg'
```

```{r, echo=FALSE}
CDE['n_op1'] <- n_op1
CDE['n_op2'] <- n_op2
write.csv(CDE, "/Users/senosam/Documents/Massion_lab/CyTOF_summary/CDE_TMA36_2021JAN13_SA_MF.csv", row.names = FALSE)
```